---
title: "EDS241: Assignment 2"
author: "Desik Somasundaram"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
--- 
  
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
packages=c("stargazer", "here", "tidyverse","stringr", "janitor", 
           "cowplot", "ggplot2", "tinytex", "datasets", "tibble", "readxl","estimatr", "car", "huxtable")

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

#devtools::install_github('rstudio/rmarkdown')
options(scipen=10) # not scientific notation


```

\noindent The goal of this assignment is to provide a simple test of whether the effects of air quality
regulations are the same across locations with different racial mix. To this end you will test if the
NOx Budget Program, a cap-and-trade market for nitrogen oxides (NOx) emissions from power
plants lead to similar effects in counties that are predominantly white versus counties that are
predominantly African American. The data are a subset of data sample I used in the following
paper: https://olivierdeschenes.weebly.com/uploads/1/3/5/0/135068654/defensive-investmentsand-the-demans-for-air-quality.pdf. You can also get more information on the NOx Budget
Program, here: https://www.epa.gov/airmarkets/nox-budget-trading-program
The data included in the file NBP.xls, which is available on Gauchospace, are: fips (fips code
identifying each county), NBP (indicator =1 if the county was regulated under the NOx Budget
Program), PctBlack (fraction of the county population that is African American), and
Dnox_masstons (change in annual NOx emissions from all power plants in a county between
2000 and 2008 (in tons)). Note that the NBP market was in effect in 212 of the 485 counties in
the sample from 2003 to 2008, so the 2008-2000 change gives us a sense of the program’s
effect on emissions. If emissions of NOx from power plants declined in a county, then
Dnox_masstons should be negative.

# Clean and plot data

\noindent The following code loads and cleans the data.

```{r , include=TRUE}

# Load data
NBPdata <- read_excel("NBP.xls", sheet = 1, na = "NA")

# Clean data
NBPdata <-janitor::clean_names(NBPdata)
```

\noindent Make a histogram depicting the distribution of dnox_masstons. 

```{r , include=TRUE}
nbphist<-ggplot(NBPdata, aes(x= NBPdata$dnox_masstons))+
  geom_histogram()+
  labs(x = "Change in annual NOx emissions from all power plants in a county between
2000 and 2008 (in tons)", y = "Count") 
```

\newpage
<center>**Figure 1: Change in annual NOx emissions from all power plants in a county between
2000 and 2008 (in tons) in CA **</center>
```{r , fig.width = 4, fig.height = 3, eval=TRUE, echo=FALSE}
nbphist
```

\noindent Figure 1 shows a skewed left distribution for the change in annual NOx emissions from all power plants in a county from 2000 to 2008.

\noindent Create an indicator = 1  if the county has PctBlack above the sample median, and = 0
otherwise (in the rest of the assignment, I refer to this variable as ‘D’). What is the average of
PctBlack for counties above the median, i.e. counties for which D=1?

```{r , include=TRUE}
#compute median percent black
pct_black_median <- median(NBPdata$pct_black)
#create indicator based on median percent black 
NBPdata<- NBPdata %>% 
  mutate(D = case_when(
    pct_black <= pct_black_median ~ 0,
    pct_black > pct_black_median ~ 1))

#find average percent black for counties above the median
NBPdata_PctBlack_avg <- NBPdata %>% 
          filter(D == 1) %>% 
          group_by(D) %>% 
          summarize(mean(pct_black))


```
The average of PctBlack for counties above the median is ***`r NBPdata_PctBlack_avg`***. 


\noindent Estimate a regression of Dnox_masstons on NBP. Interpret the estimated intercept and the
coefficient on NBP.

\begin{align}
  Y_{i} = \beta_0 + \beta_1 X_{1i} + u_{i}
\end{align}

\noindent where $Y_{i}$ is Dnox_masstons $i$, $X_{1i}$ is the NBP, and $u_{i}$ the regression error term.


```{r , include=TRUE}
model_1 <- lm_robust(formula =  dnox_masstons ~ nbp, data=NBPdata)
```

\noindent Table 1 shows the estimated coefficients from estimating equation (1). 

```{r , results = 'asis', echo = FALSE}
huxreg(model_1)
```
\noindent In model (1), the estimated $\beta_{0}$ coefficient implies that if the county was **NOT regulated** under the NOx Budget Program and if  there is a **decline of 3.622 tons** in the change in annual NOx emissions from all power plants in a county between the 2000 and 2008. The estimated $\beta_{1}$ coefficient implies that if the county was **regulated** under the NOx Budget Program there is **an additional decline of 3.92 tons** in the change in annual NOx emissions from all power plants in a county between 2000 and 2008. Both $\beta_{0}$ and $\beta_{1}$ are ***statistically significant at the 5%***.

\noindent Create an interaction between the variables NBP and D. Estimate a regression of
Dnox_masstons on NBP, D, and this interaction. Interpret each estimated regression coefficient,
including the intercept.

\begin{align}
  Y_{i} = \beta_0 + \beta_1 X_{1i} + \beta_2 X_{2i} + \beta_3 X_{1i} * X_{2i} + u_{i}
\end{align}

\noindent where $Y_{i}$ is Dnox_masstons $i$, $X_{1i}$ is the NBP, $X_{2i}$ is the D and $u_{i}$ the regression error term.


\noindent In R, we run the following code: 

```{r , include=TRUE}

model_2 <- lm_robust(formula =  dnox_masstons ~ nbp + D + nbp*D, data=NBPdata)

```

\noindent Table 2 shows the estimated coefficients from estimating equation (2). 

```{r , results = 'asis', echo = FALSE}
huxreg(model_2)
```

\noindent In model (2), the estimated $\beta_{0}$ coefficient implies that if the county was **NOT regulated** under the NOx Budget Program  **AND** the county has PctBlack **below** the sample median, there is a **decline of 2.418 tons** in the change in annual NOx emissions from all power plants in a county between the 2000 and 2008. The estimated $\beta_{1}$ coefficient implies that if the county was **regulated** under the NOx Budget
Program **AND** the county has PctBlack **below** the sample median, there is **an additional decline of 7.141 tons** in the change in annual NOx emissions from all power plants in a county between 2000 and 2008. 
The estimated $\beta_{2}$ coefficient implies that if the county was **NOT regulated** under the NOx Budget
Program **AND** the county has PctBlack **above** the sample median, there is **an additional decline of 2.588 tons** in the change in annual NOx emissions from all power plants in a county between 2000 and 2008.
The estimated $\beta_{3}$ interaction coefficient implies that if the county was **regulated** under the NOx Budget
Program **AND** the county has PctBlack **above** the sample median, the **decline** in the change in annual NOx emissions from all power plants in a county between 2000 and 2008 **gets negated by 6.372 tons**. In other words, the change in annual NOx emissions increases by 6.372 tons when both NBP and D are equal to 1.
$\beta_{0}$, $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$ are ***statistically significant at the 5%***.


\noindent What is the predicted Dnox_masstons in a county that was not regulated under NBP and
where PctBlack is above the sample median (i.e., where D=1)? Report the 95% confidence
interval for this prediction. Make sure to use “heteroskedasticity-robust” standard errors.

```{r}
PredCounty=data.frame(nbp=c(0), D=c(1))
predict(model_2, newdata=PredCounty, se.fit=TRUE, interval='confidence')
```

\noindent The predicted Dnox_masstons in a county that was not regulated under NBP and
where PctBlack is above the sample median (i.e., where D=1) is **-5 tons** with with **lower limit of -6.44 tons** and **upper limit of -3.57 tons** for the **95% confidence level**.